Sharing statistics for SPARQL Federation optimization, with emphasis on benchmark quality

### 2012-01-11 eswc-2012/phdsymp.tex
Remove stuff from fed problem
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@63 43ef1190-90fd-4dd7-a769-d0121f02b80e

An intrinsic problem of the current federated regime is that the
system first executing the federated query has no information of the
characteristics of neither the data nor the performance of the systems
it further submits queries to. This limits the techniques it can use
to optimize the query before submitting.

One aspect of this problem can be illustrated by a self-experienced
practical problem: In a federated conjuctive query, we created a query
to ask for certain books written in a certain
language. Some of the book metadata resided on one endpoint, the
language metadata was on another endpoint. Unbeknowst to us, the
distribution of languages were extremely skewed, i.e. the vast
majority of books were written in one language. Thus, sending the
language query first would in some cases be favourable, returning only
a handful of results, in other cases, it would result in a very large
number of results, which would then need to be bound in queries sent
to other endpoints, causing weak performance.

In the non-federated case, this can be mitigated by the well-known
histogram technique for subsequent selectivity estimation at query
execution time. In the federated case, this is not possible as nothing
is known about the distribution of data on individual
endpoints. \cite{5337556} addressed this problem by making it possible
to exchange histograms, but in the general case this is untenable as
the number of histograms to be computed and the number
of bins in each histogram to represent highly heterogenous data would
be very high. This would also cause the amount of statistical data to
be exchanged between endpoints to be very high.

>>>>

When using FedBench, (see \cite{Schmidt:2011:FBS:2063016.2063054}),
the authors note that the absence of statistics has very adverse
effects on performance as join order is impossible to optimize. It is
also pointed out that it causes other problems as well, for example a
high number of requests to federation members.

This problem has already received some attention, \cite{5337556}
addressed this problem by making it possible to exchange histograms,
but I suspect this will prove to become untenable as the number of
histograms to be computed and the number of bins in each histogram to
represent highly heterogenous data would be very high. This would also
cause the amount of statistical data to be exchanged between endpoints
to be very high.
